---
title: "Project_Submission_Yazan Abdullatif"
author: "Yazan Abdullatif"
format: html
editor: visual
---

Final Project submitted in completion of the bridge module Fundamentals of Programming for Digital Health WiSe 25/26\

CGM interstitial glucose data analysis\
In this project, I shall be analyzing my own interstitial glucose values, acquired by the Freestyle Libre 3 Plus continuous Glucose Monitor (CGM), over the period between 28.01.2026 and 07.02.2026. Showing the steps taken in loading, cleaning and processing the data and preparing for further utilization, as well as computing essential glycemic metrics level of the person, form which we can draw conclusions about the persons' glucose management level and condition.

```{r,include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(purrr)
library(ggplot2)
library(tidyverse)
```

```{r loading data}
# Load data
full_data <- read_csv("raw_data_small/clean_data_full.csv")
full_data$timestamp <- ymd_hms(full_data$timestamp)

# source functions
source("functions.R")
```

```{r Glycemic metrics}
# Defining glucose ranges

metrics <- calculate_metrics(full_data)
print(metrics)
```

```{r timeseries}
plot_timeseries(full_data)
```

```{r boxplot}

plot_boxplot(full_data, metrics)
```

```{r histogram}
plot_histogram(full_data, metrics)

```

```         
Summary and Conclusion
- Mean glucose: `r round(metrics$mean_glucose, 1)` mg/dL
- TIR: `r round(metrics$TIR, 1)`%
- TITR: `r round(metrics$TITR, 1)`%
- CV: `r round(metrics$cv_glucose * 100, 1)`%
```

```         
Based on the analysis, the individual demonstrates excellent glycemic control with near optimal TIR and TITR metrics
```

\
